---
title: "VPO Repertoire History"
date: "16 March 2024"
format: 
  html:
    toc: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r data}
library(tidyverse)
library(plotly)
orch <- read_lines("data/vpo-archive-performances.txt")
dates <- str_extract_all(orch, "^\\d{4}")
composers <- str_extract_all(orch, "(?<=\\t)(.*?)(?=\\:)")
piece <- str_extract_all(orch, "(?<=\\:\\ )(.*)")
piece <- lapply(piece, str_remove_all, pattern = "\\ \\([^)]*\\)")
```

```{r functions}
cat_na <- function(x) if (length(x) < 1L) NA_character_ else x
orch_tib <- tibble(
  year = list_c(lapply(dates, cat_na)),
  composer = list_c(lapply(composers, cat_na)),
  piece = list_c(lapply(piece, cat_na))
)
grab_last_name <- function(x) {
  x <- str_split(x, " ")
  lens <- lapply(x, length)
  map2_chr(x, lens, ~ .x[.y])
}
```

```{r processing}
orch_tib <- orch_tib |>
  fill(year) |>
  drop_na() |>
  mutate(across(composer:piece, ~str_squish(.x))) |>
  mutate(composer = grab_last_name(composer))

dated <- read_lines("data/vpo-rep-dated.txt") |>
  str_split(",") |>
  map(~{
    n <- length(.x)
    year = as.integer(substr(.x[n], 1, 4))
    tibble(composer = .x[1],
           piece = paste(.x[2:(n-1)], collapse = ","),
           composed = year)
  }) |>
  list_rbind() |>
  mutate(composer = grab_last_name(composer)) |>
  distinct(composer, piece, .keep_all = TRUE)

orch_tib <- left_join(orch_tib, dated) |> mutate(year = as.integer(year))
```


## Repertoire from previous seasons

* This table is a scrollable / sortable / searchable list of all pieces performed by the VPO in past seasons.
    * Click the arrows to arrange by that column
* The "composed" column was done by Chat-GPT, so may have errors or be missing.
* The date "performed" should always be correct.
* The concerto competition repertoire are most likely to be missing dates of composition
* There are some misspellings (Shoatakovich vs Shostakovich) and duplications (Symphony No. 1 vs Symphony \#1 vs Symphony \# 1)

```{r browse}
DT::datatable(
  orch_tib |> 
    relocate(composer, piece, year) |> 
    arrange(composer, piece, year) |>
    rename(performed = year),
  options = list(scrollY = "300px", paging = FALSE),
  rownames = NULL
)
```

## Visualizing previous seasons

* This shows all the music that has a date of composition. So typos and similar won't appear.
* Colours correspond to composer (though there are lots of composers)
* If you hover over a point, the composition will display.
* You can also Zoom and Pan the graphic
* Lots of Romantic music, especially this season 
    * Everything is between Nabucco (1842) and Sibelius #2 (1902), except "The Sea and Bells", the Chinese Concerti, and Prokofiev Piano Concerto

```{r show-rep}
#| column: page-inset-left
p <- orch_tib |>
  mutate(
    lab = paste(composer, piece, sep = ": "),
    alp = year <= 2014
  ) |>
  ggplot(aes(composed, year, color = composer, text = lab, alpha = alp)) +
  geom_point(shape = 16, size = 2) +
  scale_alpha_manual(guide = "none", values = c(.4, 1)) +
  scale_color_viridis_d(guide = "none") +
  geom_hline(yintercept = 2014) +
  theme_bw() +
  ylab("Year performed") + xlab("Year composed") +
  theme(legend.position = "none")
ggplotly(p, tooltip = "text")
```

## Sheet music in the VPO library

* Browseable version of the Excel sheet that Mark shared.

```{r library}
libr <- read_csv(
  "data/vpo-music-library.csv", 
  col_select = c("Composer", "Title")
) |>
  rename(Piece = Title)
DT::datatable(
  libr |> arrange(Composer, Piece),
  options = list(scrollY = "300px", paging = FALSE),
  rownames = NULL
)
```
