---
title: "VPO Repertoire History"
date: "20 March 2025"
format: 
  html:
    toc: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r packages}
library(tidyverse)
library(plotly)
```

## Repertoire from previous seasons

* This table is a scrollable / sortable / searchable list of all pieces performed by the VPO in past seasons.
    * Click the arrows to arrange by that column
* The "composed" column was done by Chat-GPT, so may have errors or be missing.
* The date "performed" should always be correct.

```{r browse}
orch_tib <- read_csv("data/orch_tib.csv")
write_rds(orch_tib, "data/orch_tib.rds")
orch_tib <- read_rds("data/orch_tib.rds")
DT::datatable(
  orch_tib |> 
    relocate(composer, piece, year) |> 
    arrange(composer, piece, year) |>
    rename(performed = year),
  options = list(scrollY = "300px", paging = FALSE),
  rownames = NULL
)
```

## Visualizing previous seasons

* This shows all the music that has a date of composition. So typos and similar won't appear.
* Colours correspond to composer (though there are lots of composers)
* If you hover over a point, the composition will display.
* You can also Zoom and Pan the graphic

```{r show-rep}
#| column: page-inset-left
p <- orch_tib |>
  mutate(
    lab = paste(composer, piece, sep = ": "),
    alp = year <= 2015
  ) |>
  ggplot(aes(composed, year, color = composer, text = lab, alpha = alp)) +
  geom_point(shape = 16, size = 2) +
  scale_alpha_manual(guide = "none", values = c(.4, 1)) +
  scale_color_viridis_d(guide = "none") +
  geom_hline(yintercept = 2015) +
  theme_bw() +
  ylab("Year performed") + xlab("Year composed") +
  theme(legend.position = "none")
ggplotly(p, tooltip = "text")
```

## Sheet music in the VPO library

* Browseable version of the Excel sheet that Mark shared.

```{r library}
libr <- read_csv(
  "data/vpo-music-library.csv", 
  col_select = c("Composer", "Title")
) |>
  rename(Piece = Title)
DT::datatable(
  libr |> arrange(Composer, Piece),
  options = list(scrollY = "300px", paging = FALSE),
  rownames = NULL
)
```
